Phylogenetics and the human microbiome.

### 2014-04-16 reviews/response.txt

> J. Zhang, P. Kapli, P. Pavlidis, A. Stamatakis: "A General Species
> Delimitation Method with Applications to Phylogenetic Placements". In
> Bioinformatics (2013), 29 (22): 2869-2876.
> It would fit will into section "The role of OTUs".

It's cited now.

> Apart from that, I believe that a more critical and slightly extended
> discussion of taxonomic classifiers in the very last part of the paper
> should be included. From my point of view, this is still a mostly
> unresolved problem and the currently best classifier, which you may
> also want to cite,  still shows pretty horrible accuracy
> (http://www.plosone.org/article/info%3Adoi%2F10.1371%2Fjournal.pone.0049334)
> at the genus and family level.

Given that taxonomic classification is not always a phylogenetic
problem, I don't think that it merits its own section. I have cited your
suggestion and other recent progress in this field for readers who want
to dig deeper.

@@

> I would also appreciate if you could explicitly mention the Gartner
> hype curve in the discussion part and provide an estimate as to where
> we are on this curve at present.

I don't see how this would contribute to the scientific goals of the
paper.

@@

Thank you for pointing me to this reference.

> 9. On page 24, when citing uBiome, it might also be fair to refer the
> similar crowd-funded American Gut Project.

Thanks.

> Typos:
> P3, L40: I would rather use "species delimitation" instead of "species
> definition".

### 2014-04-16 sbreview.tex
Responded to all reviews!

It is possible to use metagenomic data as an expanded set of marker genes.
That is, one can use 16S reads that appear in the metagenome as well as reads from other ``core" genes that are expected to follow the same evolutionary path and are present in a large proportion of micro-organisms.
This is proven to be a useful strategy, and several groups have built databases of core gene families and provide programs and/or web tools to subset a metagenome to that core set of genes \citep{von2007quantitative,wu2008amphora,stark2010mltreemap,kembel2011phylogenetic,Darling2014-ux}.
However, because of the variability of gene repertoire in microbes, this core gene set may be relatively small: even the largest collection of genes in these databases only recruits around 1 percent of a metagenome.
At least some portion of the rest of the other approximately 99 percent of the metagenome can be taxonomically classified using methods reviewed by \citep{mande2012classification}.
\citet{treangen2013metamos} report shorter run times and much higher accuracy (for taxonomic classification using reads simulated from full genomes) for such metagenomic classification when reads are assembled before they are classified.

@@

An obvious application of phylogenetics is to perform taxonomic classification, as the taxonomy is at least in part defined by phylogeny.
However, comparisons of taxonomic classification programs \citep{liu2008accurate,bazinet2012comparative} have indicated that current implementations of phylogenetic methods do not perform as well as simple classifiers based on counts of DNA substrings of a given length, which are often called $k$-mer classifiers \citep{wang2007naive,rosen2008metagenome}.
For those studies and others, the conventional metrics of classification performance such as precision and recall are applied to data simulated from known and taxonomically classified genomes.
\citet{srinivasan2012bacterial} find that phylogenetic methods to do taxonomic classification can outperform composition-based techniques at least for certain taxonomic groups.
% \citep{werner2011impact} Impact of training sets on classification of high-throughput bacterial 16s rRNA gene surveys
Some authors report that a combination of composition-based and homology-based classifiers work best \citep{brady2009phymm,parks2011classifying}.
The MEGAN program \citep{huson2007megan,huson2011integrative} BLASTs an unknown sequence onto a database of sequences with taxonomic labels and assigns the sequence the lowest (i.e.\ narrowest) taxonomic group shared by all of the high-quality hits; as such it is somewhat phylogenetic in that it uses the structure of the phylogenetic tree.
\citet{munch2008statistical,munch2008fast} infer taxonomic assignment by automatically retrieving sequences equipped with taxonomic information and building a tree on them along with an unknown sequence.
\citet{segata2012metagenomic} propose a clever approach to inferring organisms present in a metagenomic sample by compiling a database of clade-specific genes, then classifying a given read as being from the only clade that has the corresponding gene.
They show that this has good sensitivity and specificity, however, this method can only be used to identify the presence of organisms whose genome has been sequenced.

>>>>

It is possible to use metagenomic data as an expanded set of marker genes.
That is, one can use 16S reads that appear in the metagenome as well as reads from other ``core" genes that are expected to follow the same evolutionary path and are present in a large proportion of micro-organisms.
This is proven to be a useful strategy, and several groups have built databases of core gene families and provide programs and/or web tools to subset a metagenome to that core set of genes \citep{von2007quantitative,wu2008amphora,stark2010mltreemap,kembel2011phylogenetic,Darling2014-ux}.
However, because of the variability of gene repertoire in microbes, this core gene set may be relatively small: even the largest collection of genes in these databases only recruits around 1 percent of a metagenome.
At least some portion of the rest of the other approximately 99 percent of the metagenome can be taxonomically classified using a diversity of methods described below.

@@

An obvious application of phylogenetics is to perform taxonomic classification, as the taxonomy is at least in part defined by phylogeny.
However, comparisons of taxonomic classification programs \citep{liu2008accurate,bazinet2012comparative} have indicated that current implementations of phylogenetic methods do not perform as well as simple classifiers based on counts of DNA substrings of a given length, which are often called $k$-mer classifiers \citep{wang2007naive,rosen2008metagenome}.
For those studies and others, the conventional metrics of classification performance such as precision and recall are applied to data simulated from known and taxonomically classified genomes.
\citet{srinivasan2012bacterial} find that phylogenetic methods to do taxonomic classification can outperform composition-based techniques at least for certain taxonomic groups.
% \citep{werner2011impact} Impact of training sets on classification of high-throughput bacterial 16s rRNA gene surveys
Some authors report that a combination of composition-based and homology-based classifiers work best \citep{brady2009phymm,parks2011classifying}.
The MEGAN program \citep{huson2007megan,huson2011integrative} BLASTs an unknown sequence onto a database of sequences with taxonomic labels and assigns the sequence the lowest (i.e.\ narrowest) taxonomic group shared by all of the high-quality hits; as such it is somewhat phylogenetic in that it uses the structure of the phylogenetic tree.
\citet{munch2008statistical,munch2008fast} infer taxonomic assignment by automatically retrieving sequences equipped with taxonomic information and building a tree on them along with an unknown sequence.
\citet{segata2012metagenomic} propose a clever approach to inferring organisms present in a metagenomic sample by compiling a database of clade-specific genes, then classifying a given read as being from the only clade that has the corresponding gene.
They show that this has good sensitivity and specificity, however, this method can only be used to identify the presence of organisms whose genome has been sequenced.
Other metagenomic classification techniques have been reviewed by \citet{mande2012classification}; interesting recent progress has been made by \citet{Lanzen2012-ul}, \cite{Koslicki2013-mc}, and \cite{Droge2014-lj}.
\citet{treangen2013metamos} report shorter run times and much higher accuracy (for taxonomic classification using reads simulated from full genomes) for such metagenomic classification when reads are assembled before they are classified.
